---
title: "EDLD Winter 2020 Lab 2"
author: "Jim Wright; Claire Zhang"
date: "1/15/2020"
output: html_document
toc: true
toc_float: true
theme: united 
highlight: zenburn 
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      fig.width = 10,
                      fig.height = 10,
                      dev.args = list(bg="transparent"))

library(tidyverse)
library(here)
library(rio)
``` 


```{r load data, include = FALSE}

alcohol <- rio::import("https://github.com/fivethirtyeight/data/raw/master/alcohol-consumption/drinks.csv", 
                        setclass = "tbl_df")

head(alcohol)


``` 


# Data Exploration Histograms 

```{r plots, include=TRUE}

ggplot(alcohol, aes(beer_servings)) +
  geom_histogram(fill = "#56B4E9",
                 color = "white",
                 alpha = 0.9) +
  ggtitle("Bins = 30")

ggplot(alcohol, aes(beer_servings)) +
  geom_histogram(fill = "#56B4E9",
                 color = "white",
                 alpha = 0.9,
                 bins = 50) +
  ggtitle("bins = 50")

ggplot(alcohol, aes(beer_servings)) +
  geom_histogram(fill = "#56B4E9",
                 color = "white",
                 alpha = 0.9,
                 bins = 5) +
  ggtitle("Bins = 5")

ggplot(alcohol, aes(beer_servings)) +
  geom_histogram(fill = "#56B4E9",
                 color = "white",
                 alpha = 0.9,
                 bins = 25) +
  ggtitle("Bins = 25") +
  labs(x = "Beer Servings")
```

Jim: I believe the histogram of beer consumption with bins set at 25 is the superior option because the distribution is very positively skewed, so it provides a more clear explanation of how the lower end of the distribution varies. 

# Data Exploration Density Plots
```{r density plots, include=TRUE}
head(alcohol)

ggplot(alcohol, aes(wine_servings)) +
  geom_density()

ggplot(alcohol, aes(wine_servings)) +
  geom_density(fill = "#56B4E9",
               bw = 5) +
  ggtitle("BW = 5") +
  labs(x = "Wine Servings")

ggplot(alcohol, aes(wine_servings)) +
  geom_density(fill = "#56B4E9",
               bw = 0.5) +
  ggtitle("BW = 0.5") +
  labs(x = "Wine Servings")

ggplot(alcohol, aes(wine_servings)) +
  geom_density(fill = "#56B4E9",
               bw = 2) +
  ggtitle("BW = 2") +
  labs(x = "Wine Servings")

ggplot(alcohol, aes(wine_servings)) +
  geom_density(fill = "#56B4E9",
               bw = 10) +
  ggtitle("BW = 10") +
  labs(x = "Wine Servings")

ggplot(alcohol, aes(wine_servings)) +
  geom_density(fill = "#56B4E9",
               bw = 15) +
  ggtitle("BW = 15") +
  labs(x = "Wine Servings")

ggplot(alcohol, aes(wine_servings)) +
  geom_density(fill = "#56B4E9",
               bw = 20) +
  ggtitle("BW = 20") +
  labs(x = "Wine Servings")

```

Jim: I believe having a larger binwidth (10 - 20) is preferrable for the density plots because it more clearly displays the concentration of wine consumption at lower servings. Similar to beer consumption, wine consumption is skewed positivley, so I think it is better to provide a more clear display of where the largest display of consumption is occuring. 

## Assignment 1: explore the overall distribution of total alcohol consumed

1. Histogram
a. Play with the number of bins 
b. Settle on a number of bins and provide a brief justification in text

Claire: I noticed that the question is about total alcohol consumded so I thought maybe just focus on this column. And firstly I wanted to investigate the total alcohol consumption by countries but I figured I can only use geom_col instead of geom_histogram since I have two variables. Not sure this is the right thing to do. Still, it was a fun exploration, see graph "Top 20 Countries in Alcohol Consumption" below.
```{r, top 20 countries in alcohol consumption, include=TRUE}
c <- alcohol %>% 
  rename(total = total_litres_of_pure_alcohol) %>% 
  select(country, total) %>% 
  arrange(desc(total)) %>%
  slice(1:20) %>% 
  mutate(country = fct_reorder(country, total))

ggplot(c, aes(country, total)) + 
  geom_col(fill = "#56B4E9",color = "white", alpha = 0.9) + 
  coord_flip()+
  theme_minimal(base_size = 15)+
  labs(y = "Total Alcohol Consumption", x = "Top 20 Countries", title = "Top 20 Countries in Alcohol Consumption")
```

Claire: Then I set out to explore the geom_histogram. I played with the bins setting several times and found that if I set bins to 25, the graph looks nice as the scale of alcohol consumption is 0-15. However, as y-axis represents "count", I don't actually know how to interpret the graph except that total consumption seems to be lower in most countries and only a few countries have a total consumption at 10-15.
```{r}
ggplot(alcohol, aes(total_litres_of_pure_alcohol)) +
  geom_histogram(fill = "#56B4E9", color = "white", alpha = 0.9, bins = 25)+
  theme_minimal(base_size = 15)+
  labs(x = "Total Alcohol Consumption", title = "Total Alcohol Consumption Counts")
```

2. Density plot
a. compare the default bandwidth with a few other selections entered manually.
b. settle on a bandwidth selection, and justify why briefly in text.

Claire: The default bandwidth gave me a density plot as below and it's ugly (default bw = "nrd0"). 
```{r}
ggplot(alcohol, aes(total_litres_of_pure_alcohol)) +
  geom_density()
```

Claire: Then I tried to change some selections and played with different binwidths but with the bw getting bigger, more details get lost due to the smoothing, even I only set bw = 5, I'm not seeing it telling us correct info about the counts of total alcohol: 
```{r}
ggplot(alcohol, aes(total_litres_of_pure_alcohol)) +
  geom_density(fill = "#56B4E9", bw = 5, color = "white", alpha = 0.9)+
  theme_minimal(base_size = 15)
```

Claire: So I finally set bw at 0.1 since it tells me more details and allows me to see the highly different alcohol consumptions all over the world.
```{r}
ggplot(alcohol, aes(total_litres_of_pure_alcohol)) +
  geom_density(fill = "#56B4E9", bw = 0.1, color = "white", alpha = 0.9)+
  theme_minimal(base_size = 15)+
  labs(x = "Total Alcohol Liters Consumed in Each Country", y = "Counts", title = "193 Countries Total Alcohol Liters Counts")
```

# Data Reorganization 
```{r data reorganized, include=TRUE}
head(alcohol)

alcohol_pivot <- alcohol %>% 
  pivot_longer(beer_servings:wine_servings,
               names_to = "type",
               values_to = "servings")


ggplot(alcohol_pivot, aes(servings)) +
  geom_histogram(fill = "#A9E5C5",
                 color = "white",
                 alpha = 0.9,
                 bins = 5) +
  facet_wrap(~type)


ggplot(alcohol_pivot, aes(servings)) +
  geom_density(aes(fill = type),
               color = "white",
               alpha = 0.3,
               bw = 50) 

ggplot(alcohol_pivot, aes(servings, type)) +
  ggridges::geom_density_ridges(color = "white",
                                fill = "#A9E5C5")

```

Jim: All three plots suggest the distribution of alcohol consumption is positively skewed across all three types of alcohol: beer, spirts, and wine. At the lowest level of servings consumed, wine is the most often consumed type of alcohol, while at higher consumption rates, beer appears to be the drink of choice. 

# Reproduced Plot
```{r reproduced plot, include=TRUE}

?str_to_title

alcohol_pivot %>% 
  mutate(type = stringr::str_to_title(type)) %>% 
  arrange(desc(total_litres_of_pure_alcohol)) %>% 
  slice(1:(25*3)) %>% 
  ggplot(aes(country, servings)) +
  geom_col(fill = "#56B4E9", 
           alpha = 0.5) +
  coord_flip() +
  labs(x = "Country",
       y = "Proportion of total alcohol consumed",
       title = "Alcohol consumption by type",
       subtitle = "Top 25 countries displayed (according to total liters consumed")

alcohol_pivot %>% 
  mutate(type = stringr::str_to_title(type)) %>% 
  arrange(desc(total_litres_of_pure_alcohol)) %>% 
  slice(1:(25*3)) %>% 
  ggplot(aes(country, fill = total_litres_of_pure_alcohol)) +
  geom_bar(position = "fill", 
           alpha = 0.5) +
  coord_flip() +
  labs(x = "Country",
       y = "Proportion of total alcohol consumed",
       title = "Alcohol consumption by type",
       subtitle = "Top 25 countries displayed (according to total liters consumed")


alcohol_pivot %>% 
  mutate(type = stringr::str_to_title(type)) %>% 
  arrange(desc(total_litres_of_pure_alcohol)) %>% 
  slice(1:(25*3)) %>% 
  ggplot(aes(country, servings)) +
  geom_col(fill = "#56B4E9",
           alpha = 0.5) +
  coord_flip() +
  facet_wrap(~type) +
  labs(x = "Country",
       y = "Proportion of total alcohol consumed",
       title = "Alcohol consumption by type",
       subtitle = "Top 25 countries displayed (according to total liters consumed")
```

## Assignment 2: reorganize the data so you have a type and servings column.

1. Evaluate the distribution of servings by type using at least three different methods. Provide a brief substantive conclusion.

Claire: The first evaluation below is to explore my first question - which has the largest consumption in all these 193 countries, wine, spirit, or beer? The plot indicates that overall, people in these countries like beer the most and wine the least, although the mean differences are not statistically significant. 
```{r}
d <- alcohol %>% 
  rename(total = total_litres_of_pure_alcohol, wine = wine_servings, spirit = spirit_servings, beer = beer_servings)%>% 
  pivot_longer(beer:wine, names_to = "type", values_to = "servings") 
d %>% 
  ggplot(aes(type, servings))+
  geom_boxplot(fill = "#56B4E9", alpha = 0.9, width = 0.5)+
  theme_minimal(base_size = 15)+
  labs(x = "Alcohol Types", y = "Average Consumption", title = "Average Consumption of Wine, Spirit, and Beer in 193 Countries")
```

Claire: My second question is in the countries with lager alcohol consumption (pure alcohol liters 10 and up), what are people's favorite, wine, spirit, or beer? The plot below indicates that beer is still the most welcomed, but wine is consumed more than spirit in average. The mean differences are still not statistically significant though.
```{r}
e <- d %>% 
  arrange(desc(total)) %>% 
  filter(total >= 10)
e %>% 
  ggplot(aes(type, servings))+
  geom_boxplot(fill = "#56B4E9", alpha = 0.9, width = 0.5)+
  theme_minimal(base_size = 15)+
  labs(x = "Alcohol Types", y = "Average Consumption", title = "Average Consumption of Wine, Spirit, and Beer in Countries \n with More Than 10 Liters Pure Alcohol Consumption")
```

Claire: My third question is for each of the countries with lager alcohol consumption (pure alcohol liters >= 10), what is their favorite, wine, spirit, or beer? The plots below are interesting and informative, showing me, as a wine lover, countries with more wine lovers such as Andorra, Croatia, Denmark, France, Portugal, and Switzerland. What's your favorite? 
```{r, countries having alcohol consumed more than 10 liters}
e %>%
  ggplot(aes(type, servings, fill = type))+
  geom_col(alpha = 0.9, width = 1, color = "white")+
  theme_minimal()+
  facet_wrap(~country, ncol = 9)+
  labs(x = "Alcohol Types", y = "Average Consumption", title = "What Is Their Favorite?")
```

2. Reproduce plot and hopefully, extra credit (hooray!)
Claire: This was really really hard. Googled a lot but still experienced a painful trial and error process. A lot tricky things such as first I had to arrange by "wine" and mutate the "order" viriable then arrange by "total" to filter out the top 25 countries. It was so fun though to see the final plot working out.
I believe my code is too complicated to be good. Looking forward to learning from class next week!

```{r}
f <- alcohol %>% 
  rename(total = total_litres_of_pure_alcohol, wine = wine_servings, spirit = spirit_servings, beer = beer_servings) %>%
  mutate(sum = wine + spirit + beer) %>%
  arrange(wine) %>% 
  mutate(order = row_number()) %>%
  arrange(desc(total)) %>% 
  slice(1:25) %>%
  pivot_longer(beer:wine, names_to = "type", values_to = "servings") %>% 
  mutate(proportion = servings/sum)

f %>% 
  ggplot(aes(fct_reorder(country, order), proportion, fill = type))+
  geom_col(position = "fill", color = "white", alpha = 0.9, width = 1)+
  theme_minimal()+
  coord_flip(expand = FALSE)+
  labs(x = "Top Countries", y = "Proportion of Alcohol Consumed", title = "Alcohol Consumption by Type", subtitle = "Top 25 countries displayed (according to total liters consumed)")
```

